Real-time resolution of short-read assembly graph using ONT long reads

### 2020-09-14 plos_latex_clean.tex
more editting

%% END MACROS SECTION
\usepackage[normalem]{ulem}
%% npGraph
\newcommand{\npscarf}{$\mathtt{npScarf}$}
\newcommand{\npscarfg}{$\mathtt{npScarf\_wag}$}
\newcommand{\npreader}{$\mathtt{npReader}$}
\newcommand{\npanalysis}{$\mathtt{npAnalysis}$}
\newcommand{\npbarcode}{$\mathtt{npBarcode}$}
\newcommand{\npgraph}{$\mathtt{npGraph}$}
\newcommand{\lrscaf}{$\mathtt{npGraph}$}
\newcommand{\canu}{$\mathtt{Canu}$}
\newcommand{\unicycler}{$\mathtt{Unicycler}$}
\newcommand{\spades}{$\mathtt{SPAdes}$}
\newcommand{\hspades}{$\mathtt{hybridSPAdes}$}
\newcommand{\albacore}{$\mathtt{Albacore}$}
\newcommand{\racon}{$\mathtt{Racon}$}
\newcommand{\metrichor}{$\mathtt{Metrichor}$}
\newcommand{\minimap}{$\mathtt{minimap2}$}
\newcommand{\miniasm}{$\mathtt{miniasm}$}
\newcommand{\bwa}{$\mathtt{BWA\text{-}MEM}$}
\newcommand{\lrscarf}{$\mathtt{LRScaf}$}

@@

\subsection*{Algorithm  details}
The work flow of \npgraph{} mainly consists of 3 stages: (1) assembly graph pre-processing; (2) graph resolving and simplifying; (3) post-processing and reporting results. 
The first step is to load the assembly graph of Illumina contigs and analyze its components' property, including binning and multiplicity estimation.
The second step works on the processed graph and the long read data that can be provided in real-time by ONT sequencer. Based on the paths induced from long reads, the assembly graph will be resolved on-the-fly.
Finally, the graph is subjected to the last attempt of resolving and cleaning, as well as output the final results. The whole process can be managed by using either command-line interface or GUI.
Among three phases, only the first one must be performed prior to the MinION sequencing process in a streaming setup.
The algorithm works on the assembly graph of Illumina contigs, so the terms \emph{contigs} and \emph{nodes} if not mentioned specifically, would be used interchangeable throughout this context.
\subsubsection*{Contigs binning}
Contigs should belong to single or multiple groups, or \emph{bins}, that would represent different assembly units, \EG{} chromosome , plasmids, of different species if applying to a metagenomics dataset. 
A binning step is needed to assign membership of each contig to its corresponding group.

The first step is to cluster the big anchors (longer than $10$Kbp and in/out degree less than 2) based on their \emph{kmer} coverage.
To achieve this, we applied DBSCAN clustering algorithm \cite{Ester96adensity-based}, which utilises a customised metric function to map contigs into a one-dimensional space.  In order to 
 define a customised metric which is simple and fast to calculate, we assumed that a single long contig itself consists of a Poisson distribution of \emph{k-mers} count with the mean approximated by the contig's coverage. 
%The rationale is to approximate a coverage value of an extremely long contig (which can be split into more than 10,000 \emph{k-mers}) with a sampled mean of a Poisson distribution (of \emph{k-mers} count). 
The metric is then determined by a distance function of two Poisson distributions based on Kullback-Leibler divergence (or relative entropy) between the Poisson distribution representing each contig\cite{Kullback1951information}.

>>>>

%% END MACROS SECTION
\usepackage[normalem]{ulem}
%% npGraph
\newcommand{\npscarf}{$\mathtt{npScarf}$}
\newcommand{\npscarfg}{$\mathtt{npScarf\_wag}$}
\newcommand{\npreader}{$\mathtt{npReader}$}
\newcommand{\npanalysis}{$\mathtt{npAnalysis}$}
\newcommand{\npbarcode}{$\mathtt{npBarcode}$}
\newcommand{\npgraph}{$\mathtt{npGraph}$}
\newcommand{\canu}{$\mathtt{Canu}$}
\newcommand{\unicycler}{$\mathtt{Unicycler}$}
\newcommand{\spades}{$\mathtt{SPAdes}$}
\newcommand{\hspades}{$\mathtt{hybridSPAdes}$}
\newcommand{\albacore}{$\mathtt{Albacore}$}
\newcommand{\racon}{$\mathtt{Racon}$}
\newcommand{\metrichor}{$\mathtt{Metrichor}$}
\newcommand{\minimap}{$\mathtt{minimap2}$}
\newcommand{\miniasm}{$\mathtt{miniasm}$}
\newcommand{\bwa}{$\mathtt{BWA\text{-}MEM}$}
\newcommand{\lrscaf}{$\mathtt{LRScaf}$}

@@

\subsection*{Algorithm  details}
The work flow of \npgraph{} mainly consists of 3 stages: (1) assembly graph pre-processing; (2) graph resolving and simplifying; (3) post-processing and reporting results. 
The first step is to load the assembly graph of Illumina contigs and analyze its components' property, including binning and multiplicity estimation.
The second step works on the processed graph and the long read data that can be provided in real-time by ONT sequencer. Based on the paths induced from long reads, the assembly graph will be resolved on-the-fly.
Finally, the graph is subjected to the last attempt of resolving and cleaning, as well as output the final results. The whole process can be managed by using either command-line interface or GUI.
Among three phases, only the first one must be performed prior to the MinION sequencing process in a streaming setup.
The algorithm works on the assembly graph of Illumina contigs, so the terms \emph{contigs} and \emph{nodes} if not mentioned specifically, would be used interchangeable throughout this context.
\subsubsection*{Contigs binning}
Contigs should belong to single or multiple groups, or \emph{bins}, that would represent different assembly units, \EG{} chromosome , plasmids, of different species if applying to a metagenomics dataset. 
A binning step is needed to assign membership of each contig to its corresponding group. 
This information can aid the bridging stage later, as member contigs of a bridge are expected to appear in the same group. However, the criteria can be relaxed between closed groups given strong connecting evidence from long read alignments.

The first step in our binning procedure is to cluster the big anchors (longer than $10$Kbp and in/out degree less than 2) based on their \emph{kmer} coverage.
To achieve this, we applied DBSCAN clustering algorithm \cite{Ester96adensity-based}, which utilises a customised metric function to map contigs into a one-dimensional space.  In order to 
 define a customised metric which is simple and fast to calculate, we assumed that a single long contig itself consists of a Poisson distribution of \emph{k-mers} count with the mean approximated by the contig's coverage. 
%The rationale is to approximate a coverage value of an extremely long contig (which can be split into more than 10,000 \emph{k-mers}) with a sampled mean of a Poisson distribution (of \emph{k-mers} count). 
The metric is then determined by a distance function of two Poisson distributions based on Kullback-Leibler divergence (or relative entropy) between the Poisson distribution representing each contig\cite{Kullback1951information}.

### 2020-09-14 plos_latex_edit.tex
more editting

%% END MACROS SECTION
\usepackage[normalem]{ulem}
\newcommand\cancel[1]{\sout{#1}}
\newcommand\addnew[1]{\textcolor{red}{#1}}
%% npGraph
\newcommand{\npscarf}{$\mathtt{npScarf}$}
\newcommand{\npscarfg}{$\mathtt{npScarf\_wag}$}
\newcommand{\npreader}{$\mathtt{npReader}$}
\newcommand{\npanalysis}{$\mathtt{npAnalysis}$}
\newcommand{\npbarcode}{$\mathtt{npBarcode}$}
\newcommand{\npgraph}{$\mathtt{npGraph}$}
\newcommand{\lrscaf}{$\mathtt{npGraph}$}
\newcommand{\canu}{$\mathtt{Canu}$}
\newcommand{\unicycler}{$\mathtt{Unicycler}$}
\newcommand{\spades}{$\mathtt{SPAdes}$}
\newcommand{\hspades}{$\mathtt{hybridSPAdes}$}
\newcommand{\albacore}{$\mathtt{Albacore}$}
\newcommand{\racon}{$\mathtt{Racon}$}
\newcommand{\metrichor}{$\mathtt{Metrichor}$}
\newcommand{\minimap}{$\mathtt{minimap2}$}
\newcommand{\miniasm}{$\mathtt{miniasm}$}
\newcommand{\bwa}{$\mathtt{BWA\text{-}MEM}$}
\newcommand{\lrscarf}{$\mathtt{LRScaf}$}

@@

\subsection*{Algorithm  details}
The work flow of \npgraph{} mainly consists of 3 stages: (1) assembly graph pre-processing; (2) graph resolving and simplifying; (3) post-processing and reporting results. 
The first step is to load the assembly graph of Illumina contigs and analyze its components' property, including binning and multiplicity estimation.
The second step works on the processed graph and the long read data that can be provided in real-time by ONT sequencer. Based on the paths induced from long reads, the assembly graph will be resolved on-the-fly.
Finally, the graph is subjected to the last attempt of resolving and cleaning, as well as output the final results. The whole process can be managed by using either command-line interface or GUI.
Among three phases, only the first one must be performed prior to the MinION sequencing process in a streaming setup.
The algorithm works on the assembly graph of Illumina contigs, so the terms \emph{contigs} and \emph{nodes} if not mentioned specifically, would be used interchangeable throughout this context.
\subsubsection*{Contigs binning}
Contigs should belong to single or multiple groups, or \emph{bins}, that would represent different assembly units, \EG{} chromosome , plasmids, of different species if applying to a metagenomics dataset. 
A binning step is needed to assign membership of each contig to its corresponding group.

The first step is to cluster the big anchors (longer than $10$Kbp and in/out degree less than 2) based on their \emph{kmer} coverage.
To achieve this, we applied DBSCAN clustering algorithm \cite{Ester96adensity-based}, which utilises a customised metric function to map contigs into a one-dimensional space.  In order to 
 define a customised metric which is \cancel{sample} \addnew{simple} and fast to calculate, we assumed that a single long contig itself consists of a Poisson distribution of \emph{k-mers} count with the mean approximated by the contig's coverage. 
%The rationale is to approximate a coverage value of an extremely long contig (which can be split into more than 10,000 \emph{k-mers}) with a sampled mean of a Poisson distribution (of \emph{k-mers} count). 
The metric is then determined by a distance function of two Poisson distributions based on Kullback-Leibler divergence (or relative entropy) between the Poisson distribution representing each contig\cite{Kullback1951information}.

>>>>

%% END MACROS SECTION
\usepackage[normalem]{ulem}
\newcommand\cancel[1]{\sout{#1}}
\newcommand\addnew[1]{\textcolor{red}{#1}}
%% npGraph
\newcommand{\npscarf}{$\mathtt{npScarf}$}
\newcommand{\npscarfg}{$\mathtt{npScarf\_wag}$}
\newcommand{\npreader}{$\mathtt{npReader}$}
\newcommand{\npanalysis}{$\mathtt{npAnalysis}$}
\newcommand{\npbarcode}{$\mathtt{npBarcode}$}
\newcommand{\npgraph}{$\mathtt{npGraph}$}
\newcommand{\canu}{$\mathtt{Canu}$}
\newcommand{\unicycler}{$\mathtt{Unicycler}$}
\newcommand{\spades}{$\mathtt{SPAdes}$}
\newcommand{\hspades}{$\mathtt{hybridSPAdes}$}
\newcommand{\albacore}{$\mathtt{Albacore}$}
\newcommand{\racon}{$\mathtt{Racon}$}
\newcommand{\metrichor}{$\mathtt{Metrichor}$}
\newcommand{\minimap}{$\mathtt{minimap2}$}
\newcommand{\miniasm}{$\mathtt{miniasm}$}
\newcommand{\bwa}{$\mathtt{BWA\text{-}MEM}$}
\newcommand{\lrscaf}{$\mathtt{LRScaf}$}

@@

\subsection*{Algorithm  details}
The work flow of \npgraph{} mainly consists of 3 stages: (1) assembly graph pre-processing; (2) graph resolving and simplifying; (3) post-processing and reporting results. 
The first step is to load the assembly graph of Illumina contigs and analyze its components' property, including binning and multiplicity estimation.
The second step works on the processed graph and the long read data that can be provided in real-time by ONT sequencer. Based on the paths induced from long reads, the assembly graph will be resolved on-the-fly.
Finally, the graph is subjected to the last attempt of resolving and cleaning, as well as output the final results. The whole process can be managed by using either command-line interface or GUI.
Among three phases, only the first one must be performed prior to the MinION sequencing process in a streaming setup.
The algorithm works on the assembly graph of Illumina contigs, so the terms \emph{contigs} and \emph{nodes} if not mentioned specifically, would be used interchangeable throughout this context.
\subsubsection*{Contigs binning}
Contigs should belong to single or multiple groups, or \emph{bins}, that would represent different assembly units, \EG{} chromosome , plasmids, of different species if applying to a metagenomics dataset. 
A binning step is needed to assign membership of each contig to its corresponding group. 
\addnew{This information can aid the bridging phase later, as member contigs of a bridge are expected to appear in the same group. However, the criteria can be relaxed between closed groups given strong connecting evidence from long read alignments.}

The first step \addnew{in our binning procedure} is to cluster the big anchors (longer than $10$Kbp and in/out degree less than 2) based on their \emph{kmer} coverage.
To achieve this, we applied DBSCAN clustering algorithm \cite{Ester96adensity-based}, which utilises a customised metric function to map contigs into a one-dimensional space.  In order to 
 define a customised metric which is \cancel{sample} \addnew{simple} and fast to calculate, we assumed that a single long contig itself consists of a Poisson distribution of \emph{k-mers} count with the mean approximated by the contig's coverage. 
%The rationale is to approximate a coverage value of an extremely long contig (which can be split into more than 10,000 \emph{k-mers}) with a sampled mean of a Poisson distribution (of \emph{k-mers} count). 
The metric is then determined by a distance function of two Poisson distributions based on Kullback-Leibler divergence (or relative entropy) between the Poisson distribution representing each contig\cite{Kullback1951information}.
